---
title: "Quick R Question"
author: "R. Noah Padgett"
date: "5/26/2021"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

1. Using the "priority_data" file, use the variables 'coll_m06_ffq', and all '*_stool' variables to identify how many patients (participant_id) provided data for the FFQ and any stool timepoint (the codebook is "Copy of ColoCare_PriorityVar_Sharepoint_v2.4_20210310")

2. Using the "map_minimal" file, and ColocareID var, use the variable 'timepoint' to find out how many patients provided 

    i. Only one stool at baseline 
    
    ii. Both stools at baseline and at 6 months

# Packages

```{r packages, warning=F, error=F, message=F}

packages <- c(
  "tidyverse", "readxl", "kableExtra"
)
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
# Load packages
lapply(packages, library, character.only = TRUE)

w.d <- getwd()

```

# Data

```{r data, warning=FALSE, error=F, message=F}
# Priority data
priority_data <- read_xlsx("LGreathouse_Priority_Data.xlsx") 
# mapping data
map_min <- read.table("16S data/map_minimal.txt", header=T)
```

# 1. FFQ and Stool Data

1. Using the "priority_data" file, use the variables 'coll_m06_ffq', and all '*_stool' variables to identify how many patients (participant_id) provided data for the FFQ and any stool timepoint (the codebook is "Copy of ColoCare_PriorityVar_Sharepoint_v2.4_20210310")


The code below finds the number of responses per variable.

```{r}
myVar <- c("coll_m06_ffq", grep("_stool$", colnames(priority_data),value = T))


NnN <- list(
  N.Response = ~ n() - sum(is.na(.x)), 
  N.NA = ~ sum(is.na(.x))
)
A <- priority_data %>%
  filter(Module == "participant_overview") %>%
  select(all_of(myVar)) %>%
  mutate(coll_m06_ffq = as.numeric(coll_m06_ffq)) %>%
  summarise(across(everything(),NnN)) %>%
  pivot_longer(
    cols=everything(),
    names_to = "output",
    values_to = "count"
  )

A <- data.frame(
  variable = myVar,
  N.Respond = A[seq(1,9, 2),2], 
  N.NA = A[seq(2, 10, 2),2])
colnames(A) <- c("variable", "N.Respond", "N.NA")
kable(A, format="html") %>% kable_styling(full_width = T)
```

## Responses to FFQ and Stool Data

```{r}

f <- list(x = ~ifelse(is.na(.x), 0, 1))
dat <- priority_data %>%
  filter(Module == "participant_overview") %>%
  select(all_of(myVar), participant_id) %>%
  mutate(across(all_of(myVar), f)) %>%
  mutate(
    any.stool = coll_baseline_stool_x+coll_m06_stool_x+coll_m12_stool_x+coll_m24_stool_x,
    any.stool = ifelse(any.stool > 0, "Has some stool data", "Has NO stool data"),
    FFQ = ifelse(coll_m06_ffq_x == 1, "Has FFQ Data", "Does NOT have FFQ Data")
  ) %>%
  group_by(FFQ, any.stool) %>%
  summarise(Frequency=n())

kable(dat, format="html") %>%
  kable_styling(full_width = T)

```

So, __338__ patients have recorded data for the FFQ and a stool measurement at any timepoint.

# 2. Timepoint and ColocareID mapping

2. Using the "map_minimal" file, and ColocareID var, use the variable 'timepoint' to find out how many patients provided 

## i. Only one stool at baseline 
    
```{r}
map_min %>%
  filter(Timepoint == "Baseline")%>%
  summarise(N = length(unique(ColocareID)))
```

## ii. Both stools at baseline and at 6 months
    
```{r}
map_min %>%
  filter(Timepoint %in% c("Baseline", "6.Month"))%>%
  group_by(ColocareID) %>%
  mutate(N = n()) %>%
  ungroup() %>%
  filter(N==2)%>%
  summarise(N = length(unique(ColocareID)))
```

Therefore, __112__ patients have data at baseline and 6.months.


# FFQ, Baseline and 6months

```{r}
# rename id variable
map_min <- map_min %>%
  mutate(participant_id = ColocareID)
# merge into 1 data.frame
p.dat <- priority_data %>% filter(Module == "participant_overview")
merge_dat <- full_join(map_min, p.dat)
# compute identifier (Stool) for whether the patient has baseline and 6.Month 16S data
dat <- merge_dat %>%
  mutate(ffq.na = ifelse(is.na(coll_m06_ffq), 0, 1))%>%
  select(participant_id, Timepoint, ffq.na) %>%
  group_by(participant_id) %>%
  mutate(N = n()) %>%
  ungroup()%>%
  mutate(
    Stool = ifelse(N==2, "Has Stool samples at Baseline & 6.Months", "Does NOT have Stool samples at Baseline & 6.Months"),
    FFQ = ifelse(ffq.na == 1, "Has FFQ Data", "Does NOT have FFQ Data")) %>%
  group_by(Stool, FFQ) %>%
  summarise(Frequency = length(unique(participant_id)))


kable(dat, format="html") %>%
  kable_styling(full_width = T)
```

Therefore, only __2__ patients have FFQ and 16S Stool data at baseline and 6.months.


# Question 2.

## Part a. 

Of those people with stool data at baseline and/or 6 months, how many have recurrence data (b_recurrence_new_mets_yn)

```{r}

table(priority_data$Module, priority_data$b_recurrence_new_mets_yn)

table(priority_data$Module, priority_data$coll_baseline_stool)

```


```{r}
myVar <- c("coll_m06_ffq", grep("_stool$", colnames(priority_data), value = T))

# need to merge index_cancer_overview with participant overview
dat1 <- priority_data %>%
  filter(Module == "participant_overview") %>%
  select(all_of(myVar), participant_id)
dat2 <- priority_data %>%
  filter(Module == "index_cancer_overview") %>%
  select(participant_id, b_recurrence_new_mets_yn)
dat <- full_join(dat1, dat2)



f <- list(x = ~ifelse(is.na(.x), 0, 1))
dat <- dat %>%
  mutate(
    across(all_of(myVar), f),
    b6.stool = coll_baseline_stool_x+coll_m06_stool_x,
    b6.stool = ifelse(b6.stool > 0, "Has Stool samples at Baseline & 6.Months", "Does NOT have Stool samples at Baseline & 6.Months"),
  )

dat$b_recurrence_new_mets_yn[is.na(dat$b_recurrence_new_mets_yn)==TRUE] <- -99

dat$Recurr <- factor(dat$b_recurrence_new_mets_yn, levels=c(0,1,99,-99),
                     labels=c("b_recurrence_new_mets_yn==0",
                              "b_recurrence_new_mets_yn==1",
                              "b_recurrence_new_mets_yn==99",
                              "b_recurrence_new_mets_yn IS MISSING"))
  
dat.sum <- dat %>%
  group_by(Recurr, b6.stool) %>%
  summarise(Frequency=n())

kable(dat.sum, format="html") %>%
  kable_styling(full_width = T)

```


## Part b.

Of those people with stool data at baseline and/or 6 months, how many died (o_vital_status)

```{r}

table(priority_data$Module, priority_data$o_vital_status)

```


```{r}
myVar <- c("coll_m06_ffq", grep("_stool$", colnames(priority_data), value = T))

# need to merge index_cancer_overview with participant overview
dat1 <- priority_data %>%
  filter(Module == "participant_overview") %>%
  select(all_of(myVar), participant_id)
dat2 <- priority_data %>%
  filter(Module == "mr_participant_overview") %>%
  select(participant_id, o_vital_status)
dat <- full_join(dat1, dat2)



f <- list(x = ~ifelse(is.na(.x), 0, 1))
dat <- dat %>%
  mutate(
    across(all_of(myVar), f),
    b6.stool = coll_baseline_stool_x+coll_m06_stool_x,
    b6.stool = ifelse(b6.stool > 0, "Has Stool samples at Baseline & 6.Months", "Does NOT have Stool samples at Baseline & 6.Months"),
  )

dat$o_vital_status[is.na(dat$o_vital_status)==TRUE] <- -99

dat$Vital <- factor(dat$o_vital_status, levels=c(0,1,-99),
                     labels=c("Has died (o_vital_status==0)",
                              "Has NOT died (o_vital_status==1)",
                              "MISSING"))
  
dat.sum <- dat %>%
  group_by(Vital, b6.stool) %>%
  summarise(Frequency=n())

kable(dat.sum, format="html") %>%
  kable_styling(full_width = T)

```